---
title: "Ember Distance Modelling"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Ember Distance Modelling}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Model Background

The ember distance model encapsulates a statistical relationship built on top of ember simulation modelling. Fire Dynamics Simulator was user to model ember dispersion from simulated fires in Euclyptus forests under a range of weather and fuel load conditions. Statistical modelling found a simple relationship could be established between Forest Fire Danger Index (FFDI), fuel load in tonnes per hectare, and the ember distance. Embers fall in a continuous distribution away from the fire front. Our model uses the furthest distance at which an ember density of 5 embers per square meter is reached as the response variable.

The *ffdi* function can be used to calcualte the forest fire danger index for given meteorological variables, comprising temperature, relative humidity, wind speed, and drought factor.

```{r setup}
library(ggplot2)
library(fiRetools)
library(sf)
library(dplyr)

FFDI <- ffdi(Temperature=32, Humid=18,Wind=25, DF=9)
print(FFDI)
```

Using the FFDI and the fuel load (here, 12.5 tonnes/ha), we can then generate an ember transport distance using the *ember_dist* function.

```{r}
ed <- ember_dist(FFDI = FFDI,FL = 12.5)
print(ed)
```

The next step is to project the *shadow* of the polygon in the expected prevailing wind direction by that distance, incorporating some off-track ember spread, which our modelling estimated at 6.5 degrees. Load a polygon containing the boundary of the vegetation patch of interest (which is assumed to have the fuel load defined above) and use the *shift_poly* function to project it at an angle. The angle is the direction it is moving towards, and if we assume a north-west wind, this will mean a south-east direction, or 135 degrees. Polygon data should be in a projected coordinate system with metres as the spatial unit.

```{r}
veg <- read_sf("dry_forest_patch.gpkg")
sh_poly <- shift_poly(poly=veg,dis = ed,base_ang = 135, var_ang = 6.5)
```

This output polygon encompasses the original vegetation polygon, but extends it a distance and direction containing the shadow of the embers out to the distance. The *shift_poly* function can be run for each vegetation type (which may have differnet fuel loads), and then polygons merged to produce an overall ember range.

```{r}
veg_1 <- read_sf("dry_forest_patch.gpkg")
ed_1 <- ember_dist(FFDI = FFDI,FL = 12.5)
sh_poly_1 <- shift_poly(poly=veg_1,dis = ed_1,base_ang = 135, var_ang = 6.5)

veg_2 <- read_sf("wet_forest_patch.gpkg")
ed_2 <- ember_dist(FFDI = FFDI,FL = 18.5)
sh_poly_2 <- shift_poly(poly=veg_2,dis = ed_2,base_ang = 135, var_ang = 6.5)

ember_shadow <- bind_rows(sh_poly_1,sh_poly_2) %>% summarise()

```

Output ember shadow polygon can then be plotted or saved to disk.

```{r}
ggplot() + geom_sf(data=ember_shadow,fill="black") +
  geom_sf(data=veg_1,fill="green",color=NA) +
  geom_sf(data=veg_2,fill="darkgreen",color=NA)

# Save output to Geopackage
#write_sf(ember_shadow,"ember.gpkg")
```
